---
title: "CODGames"
author: "Brian Ellis"
date: "2023-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter

```{r}
rm(list = ls())
# Load libraries
library(tidyverse)
library(readxl)
```


## Task 1

1. Read the dataset into R.
•	We have not covered methods for reading an .xlsx file into R. 
•	Yes, you could save as .csv but that risks potential data loss. 
•	It would be better to search for a method (R function) that allows you to open .xlsx files in R.

```{r}
# Read dataset
COD <- read_excel("~/Downloads/CODGames2_mp.xlsx")
head(COD)
```

## Task 2- Explore 6 variables

2. Begin to explore the dataset. Explore 6 variables of your choice. (You should explore at least one categorical and at least one quantitative variable.)
•	For each, you should identify the variable, the type of variable (categorical, quantitative, etc.), information about the types of values the variable can take, and the amount of missing data within the variable. 
•	For each variable, report some summary statistics and make a visualization.
o	If a variable is categorical, make a table of possible values and include counts or proportions. Then, make an appropriate plot to show the distribution of values.
o	If a variable is quantitative, include some basic summary stats like the mean, median, standard deviation, min, max, etc. and make an appropriate plot to show the distribution of values.
•	For each variable, you should write a few sentences to describe what you have learned about the variable. Remember, the goal is not to make the plot/calculate the summaries. The goal is to use plots and summaries to gain knowledge from the data.

The first variable I will explore is the 'Result' variable.
```{r}
# List variable and data types
glimpse(head(COD$Result))
sum(is.na(COD$Result))
```

'Result' is a categorical variable even though it mainly contains integers. The values of the variable contain two sets of integer values, one the winning teams score (usually 100) and the other the losing team's score. This variable contains 0 missing values.

```{r}
# Table of possible values of 'Results'
table(COD$Result)
```

```{r}
# Visualization of 'Results'
COD %>% 
  ggplot(aes(x = Result)) +
  geom_bar()
```

What I've learned about 'Result' from doing this exploratory data analysis is that it should really be classified as a quantitative variable or at least condensed in some way. It's not that useful of a variable in its current format since at least one of the scores has to be 100. What would be nice to do would be to make a new variable out of 'Result' and calculate the difference between the winning and losing team's scores as that would allow you to get more information than the variable in its current form.

The second variable I will explore is 'Score'
```{r}
# List variable and data types
glimpse(head(COD$Score))
sum(is.na(COD$Score))
```

'Score' is a quantitative variable and contains 0 missing values. The values of the variable indicate the score from a match which is probably dependent on several other variables from this dataset.

```{r}
# Include summary statistics
summary(COD$Score)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Score)) +
  geom_histogram(bins = 20, fill = "dodgerblue", color = "grey") +
  labs(title = "Distribution of Scores",
       x = "Score",
       y = "Frequency")
```

From doing this initial exploration of the 'Score' variable, we can see that it is a continuous quantitative variable that ranges from around 100-7000. Both the median and the mode lay just under 3000 so that seems like the average value of score. Looking at the histogram plotted, the distribution of scores seems to be a little right-skewed which would mean that infrequent high scores are possible. 