---
title: "CODGames"
author: "Brian Ellis"
date: "2023-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter

```{r}
rm(list = ls())
# Load libraries
library(tidyverse)
library(readxl)
```


## Task 1

1. Read the dataset into R.
•	We have not covered methods for reading an .xlsx file into R. 
•	Yes, you could save as .csv but that risks potential data loss. 
•	It would be better to search for a method (R function) that allows you to open .xlsx files in R.

```{r}
# Read dataset
COD <- read_excel("~/Downloads/CODGames2_mp.xlsx")
head(COD)
```

## Task 2- Explore 6 variables

2. Begin to explore the dataset. Explore 6 variables of your choice. (You should explore at least one categorical and at least one quantitative variable.)
•	For each, you should identify the variable, the type of variable (categorical, quantitative, etc.), information about the types of values the variable can take, and the amount of missing data within the variable. 
•	For each variable, report some summary statistics and make a visualization.
o	If a variable is categorical, make a table of possible values and include counts or proportions. Then, make an appropriate plot to show the distribution of values.
o	If a variable is quantitative, include some basic summary stats like the mean, median, standard deviation, min, max, etc. and make an appropriate plot to show the distribution of values.
•	For each variable, you should write a few sentences to describe what you have learned about the variable. Remember, the goal is not to make the plot/calculate the summaries. The goal is to use plots and summaries to gain knowledge from the data.

The first variable I will explore is the 'Result' variable.
```{r}
# List variable and data types
glimpse(head(COD$Result))
sum(is.na(COD$Result))
```

'Result' is a categorical variable even though it mainly contains integers. The values of the variable contain two sets of integer values, one the winning teams score (usually 100) and the other the losing team's score. This variable contains 0 missing values.

```{r}
# Table of possible values of 'Results'
table(COD$Result)
```

```{r}
# Visualization of 'Results'
COD %>% 
  ggplot(aes(x = Result)) +
  geom_bar()
```

What I've learned about 'Result' from doing this exploratory data analysis is that it should really be classified as a quantitative variable or at least condensed in some way. It's not that useful of a variable in its current format since at least one of the scores has to be 100. What would be nice to do would be to make a new variable out of 'Result' and calculate the difference between the winning and losing team's scores as that would allow you to get more information than the variable in its current form.

The second variable I will explore is 'Score'
```{r}
# List variable and data types
glimpse(head(COD$Score))
sum(is.na(COD$Score))
```

'Score' is a quantitative variable and contains 0 missing values. The values of the variable indicate the score from a match which is probably dependent on several other variables from this dataset.

```{r}
# Include summary statistics
summary(COD$Score)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Score)) +
  geom_histogram(bins = 20, fill = "dodgerblue", color = "grey") +
  labs(title = "Distribution of Scores",
       x = "Score",
       y = "Count")
```

From doing this initial exploration of the 'Score' variable, we can see that it is a continuous quantitative variable that ranges from around 100-7000. Both the median and the mode lay just under 3000 so that seems like the average value of score. Looking at the histogram plotted, the distribution of scores seems to be a little right-skewed which would mean that infrequent high scores are possible. 

The third variable I will explore is 'PrimaryWeapon'
```{r}
# List variable and data types
glimpse(head(COD$PrimaryWeapon))
sum(is.na(COD$PrimaryWeapon))
```

The variable 'PrimaryWeapon' is a categorical variable with 0 missing values. It looks like this variable lists the primary gun used by the player in a match. It would be interesting to see the effect this variable has on several factors such as 'Score'.

```{r}
# Table of possible values of 'Results'
table(COD$PrimaryWeapon)
```

```{r}
# Visualize distribution
COD %>% 
  ggplot(aes(x = PrimaryWeapon)) +
  geom_bar(aes(fill = PrimaryWeapon)) +
  labs(title = "Count of Primary Weapons",
       x = "Type of Weapon",
       y = "Count")
```

After looking at this table and visualization, it seems like there are a few favorite weapons of the players in this dataset. There are about 4 weapons that dominate the share of the distribution while the others show up infrequently. It would be nice to correlate the scores of players with their weapon choice so as to justify the use of a certain kind of weapon.


The fourth variable I will explore is 'Eliminations'
```{r}
# List variable and data types
glimpse(head(COD$Eliminations))
sum(is.na(COD$Eliminations))
```

The variable is quantitative and has 0 missing values. It looks like it is the number of total kills a player has in a match. 

```{r}
# Include summary statistics
summary(COD$Eliminations)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Eliminations)) +
  geom_histogram(binwidth = 3, fill = "violet", color = "grey") +
  labs(title = "Distribution of Eliminations",
       x = "Eliminations",
       y = "Count")
```

This variable has a similar distribution and spread to the 'Score' variable which leads me to believe these two variables are highly correlated. We see the majority of elimination totals around 15 with a few outliers in the higher range.

```{r}
cor(COD$Score, COD$Eliminations)
```

We can confirm the correlation above, the two variables have a correlation of 0.836.

The fifth variable I will explore is 'Deaths'
```{r}
# List variable and data types
glimpse(head(COD$Deaths))
sum(is.na(COD$Deaths))
```

The variable is quantitative and has 0 missing values. It looks like it is the number of total number of deaths a player has in a match. 

```{r}
# Include summary statistics
summary(COD$Deaths)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Deaths)) +
  geom_histogram(binwidth = 3, fill = "black", color = "grey") +
  labs(title = "Distribution of Deaths",
       x = "Deaths",
       y = "Count")
```

'Deaths' seems to be normally distributed with a mean and median hovering around 15. Just like 'Eliminations', I feel this variable is highly correlated with 'Score' and it would be interesting to plot their relationships.

The sixth variable I will explore is 'Damage'
```{r}
# List variable and data types
glimpse(head(COD$Damage))
sum(is.na(COD$Damage))
```

'Damage' is a quantitative variable with 0 missing values. The values of this variable seem to be the damage a player takes or inflicts in a match.

```{r}
# Include summary statistics
summary(COD$Damage)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Damage)) +
  geom_histogram(fill = "red", color = "grey") +
  labs(title = "Distribution of Damage",
       x = "Damage",
       y = "Count")
```

'Damage' seems to be normally distributed with an average around 400. I would be interested to see the relationship this variable shares with 'PrimaryWeapon' as to see if a particular weaopn leads to higher damage.

## Task 3

Begin a list of questions that you have about the dataset. Suppose you are working on a project and your manager gives you this dataset. Write 3 questions that you would ask your manager.
•	You have not been given any information about what the variables represent. So, if you are unsure about the meaning of any of the variables, write your questions.
•	You have not been given any information about how the data were collected. If you have questions about the data collection, write them down.
•	If you have any other questions, write them down.
•	NOTE: Even if you are familiar with the game and think that you understand the meaning of the variables, you are still expected to write 3 potential questions that you could ask.