---
title: "CODGames"
author: "Brian Ellis"
date: "2023-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Front Matter

```{r}
rm(list = ls())
# Load libraries
library(tidyverse)
library(readxl)
library(lubridate)
```


## Task 1

```{r}
# Read dataset
COD <- read_excel("~/Downloads/CODGames2_mp.xlsx")
head(COD)
```

## Task 2- Explore 6 variables

The first variable I will explore is the 'Result' variable.
```{r}
# List variable and data types
glimpse(head(COD$Result))
sum(is.na(COD$Result))
```

'Result' is a categorical variable even though it mainly contains integers. The values of the variable contain two sets of integer values, one the winning teams score (usually 100) and the other the losing team's score. This variable contains 0 missing values.

```{r}
# Table of possible values of 'Results'
table(COD$Result)
```

```{r}
# Visualization of 'Results'
COD %>% 
  head(50) %>%  # I use the head function here because plotting the entire dataset is a visual mess and doesn't add any value
  ggplot(aes(x = Result)) +
  geom_bar() +
  labs(title = "Count of Results",
       x = "Result",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

What I've learned about 'Result' from doing this exploratory data analysis is that it should really be classified as a quantitative variable or at least condensed in some way. It's not that useful of a variable in its current format since at least one of the scores has to be 100. What would be nice to do would be to make a new variable out of 'Result' and calculate the difference between the winning and losing team's scores as that would allow you to retrieve more information as compared to the variable in its current form.

The second variable I will explore is 'Score'
```{r}
# List variable and data types
glimpse(head(COD$Score))
sum(is.na(COD$Score))
```

'Score' is a quantitative variable and contains 0 missing values. The values of the variable indicate the score from a match which is probably dependent on several other variables from this dataset.

```{r}
# Include summary statistics
summary(COD$Score)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Score)) +
  geom_histogram(bins = 20, fill = "dodgerblue", color = "grey") +
  labs(title = "Distribution of Scores",
       x = "Score",
       y = "Count")
```

From doing this initial exploration of the 'Score' variable, we can see that it is a continuous quantitative variable that ranges from around 100-7000. Both the median and the mode lay just under 3000 so that seems like the average value of score. Looking at the histogram plotted, the distribution of scores seems to be a little right-skewed.

The third variable I will explore is 'PrimaryWeapon'
```{r}
# List variable and data types
glimpse(head(COD$PrimaryWeapon))
sum(is.na(COD$PrimaryWeapon))
```

The variable 'PrimaryWeapon' is a categorical variable with 0 missing values. It looks like this variable lists the primary gun used by the player in a match. It would be interesting to see the effect this variable has on several factors such as 'Score'.

```{r}
# Table of possible values of 'PrimaryWeapon'
table(COD$PrimaryWeapon)
```

```{r}
# Visualize distribution
COD %>% 
  ggplot(aes(x = PrimaryWeapon)) +
  geom_bar(aes(fill = PrimaryWeapon)) +
  labs(title = "Count of Primary Weapons",
       x = "Type of Weapon",
       y = "Count")
```

After looking at this table and visualization, it seems like there are a few favorite weapons of the players in this dataset. There are about 4 weapons that dominate the share of the distribution while the others show up infrequently. It would be nice to correlate the scores of players with their weapon choice so as to justify the use of a certain kind of weapon.


The fourth variable I will explore is 'Eliminations'
```{r}
# List variable and data types
glimpse(head(COD$Eliminations))
sum(is.na(COD$Eliminations))
```

The variable is quantitative and has 0 missing values. It looks like it is the number of total kills a player has in a match. 

```{r}
# Include summary statistics
summary(COD$Eliminations)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Eliminations)) +
  geom_histogram(binwidth = 3, fill = "violet", color = "grey") +
  labs(title = "Distribution of Eliminations",
       x = "Eliminations",
       y = "Count")
```

This variable has a similar distribution and spread to the 'Score' variable which leads me to believe these two variables are highly correlated. We see the majority of elimination totals around 15 with a few outliers in the higher range.

```{r}
cor(COD$Score, COD$Eliminations)
```

We can confirm the correlation above, the two variables 'Score' and 'Eliminations' have a correlation of 0.836.

The fifth variable I will explore is 'Deaths'
```{r}
# List variable and data types
glimpse(head(COD$Deaths))
sum(is.na(COD$Deaths))
```

The variable is quantitative and has 0 missing values. It looks like it is the number of total number of deaths a player has in a match. 

```{r}
# Include summary statistics
summary(COD$Deaths)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Deaths)) +
  geom_histogram(binwidth = 3, fill = "black", color = "grey") +
  labs(title = "Distribution of Deaths",
       x = "Deaths",
       y = "Count")
```

'Deaths' seems to be normally distributed with a mean and median hovering around 15. Just like 'Eliminations', I feel this variable is highly correlated with 'Score' and it would be interesting to plot their relationships.

The sixth variable I will explore is 'Damage'
```{r}
# List variable and data types
glimpse(head(COD$Damage))
sum(is.na(COD$Damage))
```

'Damage' is a quantitative variable with 0 missing values. The values of this variable seem to be the damage a player takes or inflicts in a match.

```{r}
# Include summary statistics
summary(COD$Damage)
```


```{r}
# Visualization of variable
COD %>% 
  ggplot(aes(x = Damage)) +
  geom_histogram(fill = "red", color = "grey") +
  labs(title = "Distribution of Damage",
       x = "Damage",
       y = "Count")
```

'Damage' seems to be normally distributed with an average around 400. I would be interested to see the relationship this variable shares with 'PrimaryWeapon' as to see if a particular weapon leads to higher damage.

## Task 3

- Some of the variables are unclear in their names, like what does 'Damage' represent? The damage a player inflicts in a match or the damage that they receive?
- It seems some of the variables need to be altered in order to get meaning from them, like 'Result' for instance should be mutated so as to make the variable more useful. Could we take the difference between the two numbers to make a more informative variable?
- From where is this data sourced? Is this a collection of many different player's stats from a random time period, or a set time period? Is this one player's stats or many different player's stats?

## Task 4 Part 1

Is the player’s performance, as quantified by the amount of experience points gained (TotalXP variable) changing over time?

```{r}
# Extracting month and creating new variable
COD <-
  COD %>% 
    mutate(Month = month(as.POSIXlt(Date, format = "%Y/%m/%d"))) %>% 
  mutate(Month = month.name[Month])

```

```{r}
# Side by Side boxplots
level_order <- c('June', 'July', 'August') 

COD %>% 
  ggplot(aes(x = factor(Month, levels = level_order), y = TotalXP, color = Month)) +
  geom_boxplot() +
  labs(title = "Total Experience by Month",
       x = "Month",
       y = "Experience")
```

There doesn't seem to be a major discrepency between amount of experience gained over time as the IQRs of each boxplot overlay each other. This tells us that the player's performance as quantified by total experience is not changing over time. We should include more variables into our analysis that would give us more information as to why the total experience is not displaying significant differences over time. Variables such as amount of data points/games played in each month or type of game mode played could significantly be affecting player experience.

## Task 4 Part 2

The question I want to answer is, what immediate changes can the player make to improve their overall performance as measured by Score?

### Game Type

To do this, we will quantify performance by several variables that that have an impact on 'Score'. Let's start with the 'GameType' variable which tells us the game mode the player chooses to play in.

```{r}
COD %>% 
  ggplot(aes(x = GameType)) +
  geom_bar() +
  labs(x = "Game Type")
```

The player heavily favors Team Deathmatch, but is it their best performing game type?
```{r}
COD %>% 
  group_by(GameType) %>% 
  summarize(N = n(),
            AvgScore = sum(Score) / N,
            AvgElim = sum(Eliminations) / N) %>% 
  arrange(desc(AvgScore))
```

Their best performing game type is Domination, but with only 3 games played it doesn't give us enough confidence to be sure. One suggestion we can make is tell the player to play more Domination game types.

### Map

Now we'll focus on another area the player can control, the type of map they end up playing in. Although map selection is ultimately decided on by vote or random chance, if a map that is unfavorable towards the player's score is chosen they can leave the game session and queue for another map that is better for them. Let's look at the player's better performing maps:

```{r}
# Maps that give the most score
COD %>% 
  group_by(Choice) %>% 
  summarize(N = n(),
            TotalScore = sum(Score),
            AvgScore = TotalScore / n()) %>% 
  arrange(desc(AvgScore)) %>% 
  head(10)
```

Now the lowest performing maps:

```{r}
# Maps that give the lowest score
COD %>% 
  group_by(Choice) %>% 
  summarize(N = n(),
            TotalScore = sum(Score),
            AvgScore = TotalScore / n()) %>% 
  arrange(AvgScore) %>% 
  head(10)
```

Jungle seems to be especially bad. Let's create a linear regression model with the dependent variable 'Score' and the independent variable 'Choice' to see if we find some statistically significant results.

```{r}
lmChoice <- lm(Score ~ Choice, data = COD)
summary(lmChoice)
```

From these results, we see that the map 'Jungle' should be avoided by the player due to its significantly low score. Its p-value of less than 0.05 indicates that the poor results from the map Jungle are statistically significant enough to not be random chance. If that map ends up being chosen, the player should exit the match and queue into a different game.

### Primary Weapon

One of the most obvious changes a player can make is in the choice of their primary weapon. Let's take a closer look at this variable and the relationship it shares with 'Score'.

```{r}
COD %>% 
  ggplot(aes(x = Score, group = PrimaryWeapon, color = PrimaryWeapon, alpha = 0.1)) +
  geom_density() +
  labs(title = "Score by Primary Weapon",
       x = "Score",
       y = "Density")
```

From this density plot there doesn't seem to be a convincingly dominant weapon that would prove the best change although AK-47 seems to give the player the widest distribution and the highest frequency of better scores.

```{r}
COD %>% 
  ggplot(aes(x = PrimaryWeapon, y = Score)) +
  geom_boxplot() +
  geom_point(color = "red", position = "jitter") +
  labs(title = "Score by Primary Weapon")
```

This boxplot allows us to get a better visual of the frequency and discrepancy amongst the weapons. Although it looked like AK-47 provided the player with the highest scores on average, we can see there are very few games played with the weapon so that data can be misleading. It looks like the gun Type 63 could be a better performing gun for the player.

Now let's see if the player's choice in gun is changing over time.

```{r}
COD %>% 
  ggplot(aes(x = PrimaryWeapon)) +
  geom_bar(aes(fill = Month), position = "dodge")
```

There are several guns that were only used in a particular month, indicating the player uses one gun consistently for a certain period of time before moving onto the next gun.

```{r}
COD %>% 
  ggplot(aes(x = factor(Month, levels = level_order))) +
  geom_bar(aes(fill = PrimaryWeapon), position = "dodge")
```

If we plot the 'Month' variable on the x-axis, we see that the player is definitely changing their weapon of choice over time. Here we see a trend towards using more of the Pelington 703 and Type 63 as time goes on. Judging from the earlier boxplots, the player should keep using Type 63 but try switching the Pelington 703 for another weapon because of its poor performance, maybe trying the Ak-47 again.

Let's create a linear regression model using 'PrimaryWeapon' as a predictor for 'Score'.

```{r}
lmWeapon <- lm(Score ~ PrimaryWeapon, data = COD)
summary(lmWeapon)
```

Looking at the summary we can see a statistically significant result for the weapons AK-47 (the intercept), M16, and Pelington 703. In the case of AK-47, the model reports the highest scores and we would suggest that the player should utilize this weapon more. For weapons M16 and Pelington 703 there was significant decline in 'Score' from which we can suggest to the player that they should stop using those two weapons. The only drawback to this analysis is the lack of data points for the Ak-47, we saw in a previous boxplot the sparseness of the data concerning the AK-47 which doesn't give us a lot of confidence in our prediction that its usage will make a signficant increase in the players 'Score'. By looking at a few variables that are under the player's control, 'Choice', 'PrimaryWeapon', and 'GameType', we can suggest that the player takes the following actions to improve their overall performance as measured by 'Score': 
<ul>
<li>Do not use M16 and Pelington 703 as a primary weapon - instead, try using the Ak-47 more often</li>
<li>Play only Domination and TDM Game Types and try playing more Domination</li>
<li>Avoid matches where the map 'Jungle' is chosen</li></ul>

We can analyze the potential improvement in 'Score' if the player takes these measures:

```{r}
CODChange <-
  COD %>% 
  filter(GameType == "HC - TDM" | GameType == "HC - Domination",
         Choice != "Jungle",
         PrimaryWeapon != "M16", PrimaryWeapon != "Pelington 703") %>% 
  select(Choice, GameType, PrimaryWeapon, Eliminations, Damage, Score, TotalXP)
head(CODChange)
```

```{r}
# Create dataframe containing filtered mean score and original mean score
Compare <- data.frame(Dataset = c("Original", "Filtered"),
                        AvgScore = c(mean(COD$Score),
                      mean(CODChange$Score)))
Compare
# Find improvement in mean score between datasets
Compare %>% 
  pivot_wider(names_from = Dataset, values_from = AvgScore) %>% 
  summarize(Difference = Filtered - Original)
```

```{r}
Compare %>% 
  ggplot(aes(x = Dataset, y = AvgScore, fill = Dataset)) +
  geom_col() +
  labs(title = "Comparison of Average Score",
       x = "Dataset",
       y = "Mean of Score") +
  geom_segment(aes(x = 1.5, y = 3251, xend = 1.5, yend = 2950), color = "red", size = 1, arrow = arrow(length = unit(0.08, "inches"))) +
  geom_text(aes(x = 1.95, y = 3100), label = "Avg Score Difference of 306", color = "red")
```

By comparing the mean score of the original data and the filtered data, which contains only the suggestions we propose the player to make, we see a 'Score' improvement of 306. All of these changes are immediate, actionable steps the player can make the next time they play to increase their performance as measured by 'Score'.
